---
title: "Customize your pkgdown website aspect"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{customization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pkgdown)
```

You can alter the design/styling of your pkgdown website to suit your taste and brand. _Theming_ refers to being able to tweak colours, fonts, spacing etc. by providing only Bootstrap variables, Bootstrap rules, or any custom CSS. _Layout_ refers to being able to more dramatically change the content of say the navbar by providing your own HTML templates to pkgdown.

Another key concept related to styling your pkgdown website is whether you are tweaking _one_ package website, or several ones. 
In the latter case you might want to share your style files in the form of a package.

## Theming

### Choosing a bootswatch theme

You can choose any [Bootswatch theme](https://bootswatch.com/) that's compatible with the Bootstrap version you use.
Run either `bslib::bootswatch_themes(3)` or `bslib::bootswatch_themes(4)` to get a list.
If you were using the paper theme, note that it was renamed [materia](https://blog.bootswatch.com/post/167233472837/announcing-bootswatch-400).

## Layout

When not to use layout: if using "in_header.html", "before_body.html" or "after_body.html".
Maintaining your own "head.html" template requires more work as you might need to keep it up-to-date with upstream changes in pkgdown itself.

## Sharing pkgdown styles
