---
title: "Customise your site"
output: rmarkdown::html_vignette
description: >
  Learn how to change the look and feel of your pkgdown site.
vignette: >
  %\VignetteIndexEntry{Customise your site}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette teaches you how to customise the style/design of your pkgdown site.
We'll start by discussing two techniques that only require tweaks to your `_pkgdown.yml`: theming (colours and fonts) and layout (content of the navbar, sidebar, footer, ...).
We'll then discuss how to add additional HTML and other files.
Next, we'll discuss how to give multiple sites the same style using a package, then finish up with some workflow advice.

In terms of your `_pkgdown.yml`, this vignette focusses on the most important fields nested under `template` and `navbar`. To learn more about customising other aspects of the site, see the documentation for the indiviudal functions like `build_reference()`, `build_articles()`,  `build_home()`, `build_redirects()`, and `init_site()`. To learn about less important fields nested under `template`, see `build_site()`.

```{r setup}
library(pkgdown)
```

## Getting started

Most theming features work only with Bootstrap 5, so first update your site by adding the following lines to your `_pkgdown.yml`:

``` yaml
template:
  bootstrap: 5
```

Overall, the site should look pretty similar, but you will notice a number of small improvements.
Most importantly, the default font is much bigger, making it considerably easier to read.
Upgrading to Bootstrap 5 has a low chance of breaking your site unless you were using your [own pkgdown templates](#template-packages) or custom CSS.

## Theming

There are two ways to change the visual style of your site from `_pkgdown.yml`: using a pre-packaged bootswatch theme or customising theme variables with [bslib](https://rstudio.github.io/bslib/).
The following sections show you how.

Please note that pkgdown's default theme has been carefully optimised to be accessible, so if you make changes, make sure to also read `vignette("accessibility")` to learn about potential accessibility pitfalls.

### Light switch {#light-switch}

You can provide a "light switch" to allow your users to switch between dark and light themes by setting the `light-switch` template option to true:

```yaml
template:
  light-switch: true
```

This will add a `lightswitch` component to the navbar, which by default appears at the far right. This allows the user to select light mode, dark mode, or auto mode (which follows the system setting). The modes are applied using Bootstrap 5.3's [colours modes](https://getbootstrap.com/docs/5.3/customize/color-modes/) so are not separate themes, but a thin layer of colour customisation applied via CSS.

### Bootswatch themes

The easiest way to change the entire appearance of your website is to use a [Bootswatch theme](https://bootswatch.com):

``` yaml
template:
  bootstrap: 5
  bootswatch: materia
```

(Themes are unlikely to work with the light switch, but you can try it and see.)

Changing the bootswatch theme affects both the HTML (via the navbar, more on that below) and the CSS, so you'll need to re-build your complete site with `build_site()` to fully appreciate the changes.
While you're experimenting, you can speed things up by just rebuilding the home page and the CSS by running `build_home_index()`; `init_site()` (and then refreshing the browser).

Bootswatch templates with tall navbars (e.g. lux, pulse) also require that you set the `pkgdown-nav-height` bslib variable.
Because Bootswatch themes are provided by the [bslib](https://rstudio.github.io/bslib/) R package, you can also nest the `bootswatch` field under the `bslib` field.

``` yaml
template:
  bootstrap: 5
  bslib:
    bootswatch: lux
    pkgdown-nav-height: 100px
```

You can find the correct height by running `$(".navbar").outerHeight()` in the [javascript console](https://firefox-source-docs.mozilla.org/devtools-user/web_console/index.html).

### bslib variables

Instead of picking a complete theme, you can tweak fonts and colours individually using bslib variables.
[bslib](https://rstudio.github.io/bslib/) is an R package that wraps sass, the tool that Boostrap uses to produce CSS from a special language called [scss](https://sass-lang.com).
The primary advantage of scss over CSS is that it's more programmable, so you can have a few key bslib variables that affect appearance of many HTML elements.

There are three key variables that affect the colour:

-   `bg` (background) determines the page background.
-   `fg` (foreground) determines the text colour. `bg` and `fg` are mixed to yield `gray-100`, `gray-200`, ..., `grey-900`, which are used to style other elements to match the overall colour scheme.
-   `primary` sets the link colour and the (translucent) hover colour in the navbar and sidebar.

``` yaml
template:
  bootstrap: 5
  bslib:
    bg: "#202123"
    fg: "#B8BCC2"
    primary: "#306cc9"
```

You can customise other components by setting more specific bslib variables, taking advantage of inheritance where possible.
For example, `table-border-color` defaults to `border-color` which defaults to `gray-300`.
If you want to change the colour of all borders, you can set `border-color` and if you just want to change the colour of table borders, you can set `table-border-color`.
You can find a full list of variables in the [bslib docs](https://rstudio.github.io/bslib/articles/bs5-variables/index.html).

If you're using the light switch, [many colours](https://getbootstrap.com/docs/5.3/customize/color-modes/#sass-variables) are available for customisation specifically for the dark theme.

Theming with bslib is powered by `bslib::bs_theme()` and the `bslib` field is a direct translation of the arguments to that function.
As a result, you can fully specify a bslib theme using the `template.bslib` field, making it easy to share YAML with the `output.html_document.theme` field [of an R Markdown document](https://rstudio.github.io/bslib/articles/theming/index.html).

``` yaml
template:
  bslib:
    version: 5
    bg: "#202123"
    fg: "#B8BCC2"
    primary: "#306cc9"
```

While iterating on colours and other variables you only need to rerun `init_site()` and refresh your browser to see the changes.

### Fonts

You can also override the default fonts used for the majority of the text (`base_font`), for headings (`heading_font`) and for code (`code_font`).
The easiest way is to supply the name of a [Google font](https://fonts.google.com) with the following syntax:

``` yaml
template:
  bootstrap: 5
  bslib:
    base_font: {google: "Roboto"}
    heading_font: {google: "Roboto Slab"}
    code_font: {google: "JetBrains Mono"}
```

If you want to use a non-Google font, you'll need to do a bit more work. There are two steps: you need to first configure the font with CSS and then use it in your `_pkgdown.yml`. There are two ways you might get the CSS:

* As a block of CSS which you should put in `pkgdown/extra.scss` or `pkgdown/extra.css`. The CSS will look something like this:

    ```css
    @font-face {
      font-family: "proxima-nova";
      src: 
        local("Proxima Nova Regular"), 
        local("ProximaNova-Regular"),
        url("https://example.com/ProximaNova-Regular.eot?#iefix") format("embedded-opentype"),
        url("https://example.com/fonts/proxima/ProximaNova-Regular.woff2") format("woff2"),
        url("https://example.com/fonts/proxima/ProximaNova-Regular.woff") format("woff"),
        url("https://example.com/fonts/proxima/ProximaNova-Regular.ttf") format("truetype");
      font-weight: normal;
      font-style: normal;
      font-display: fallback;
    }
    ```

* As a link to a style file, which you'll need to add to the `<head>` using this syntax:

    ```yaml
    template:
    includes:
      in_header: <link rel="stylesheet" type="text/css" href="https://..." />
    ```

Then in `_pkgdown.yml` you can use the name of the font you just specified:

```yaml
template:
  bslib:
    base_font: proxima-nova
```

Depending on where the font is from (and if you purchased it), you may need to take additional steps to ensure that it can only be used from your site, and/or make sure that it can still be used when you're previewing locally. If you're having problems getting a custom font to work, looking for errors in the [browser developer console](https://developer.mozilla.org/en-US/docs/Glossary/Developer_Tools) is a good place to start.

When iterating on fonts, you'll need to run `build_home_index(); init_site()` then refresh your browser to see the update.


### Syntax highlighting

The colours used for syntax highlighting in code blocks are controlled by the `theme` setting:

``` yaml
template:
  bootstrap: 5
  theme: breeze-light
```

You can choose from any of the following options: `r paste0(pkgdown:::highlight_styles(), collapse = ", ")`.

Bootswatch themes with a dark background (e.g. cyborg, darkly, solar) will need a dark syntax highlighting `theme`, e.g. `arrow-dark`:

``` yaml
template:
  bootstrap: 5
  bootswatch: cyborg
  theme: arrow-dark
```

If you're using the light switch, you will want to provide a `theme` and a `theme-dark`:

```yaml
template:
  light-switch: true
  theme: gruvbox-light
  theme-dark: gruvbox-dark
```

The foreground and background colours used for inline code are controlled by `code-color` and `code-bg` bslib variables.
If you want inline code to match code blocks, you'll need to override the variables yourself, e.g.:

``` yaml
template: 
  bootstrap: 5
  theme: arrow-dark  
  bslib:
    code-bg: "#2b2b2b"
```

### Math rendering

By default, pkgdown will render math using mathml. mathml is the official standard for rendering math on the web, and requires no additional javascript or css dependencies. However, browser support for complex math is not always that good, so if you are including complex equations in your documentation, you may want to switch to either [`katex`](https://katex.org) or [`mathjax`](https://www.mathjax.org) by using the `template.math-rendering` field:

```yaml
template:
  math-rendering: katex
```

### Navbar style

The primary navbar colours are determined by HTML classes, not CSS, and can be customized using the `navbar` fields `bg` and `type` which control the background and foreground colours respectively.
Typically `bg` will be one of `light`, `dark`, or `primary`:

``` yaml
navbar:
  bg: primary
```

You generally don't need to set `bg` if you use a bootswatch theme, as pkgdown will pick the `bg` used on the [Bootswatch preview](https://bootswatch.com/).
Similarly, you don't usually need to set `type` because bootstrap will guess it for you.
If the guess is wrong, you can override with `type: light` or `type: dark` depending on whether the background colour is light (so you need dark text) or `type: dark` if the background is dark (so you need light text).
Unfortunately, these are defined relative to the page background, so if you have a dark site you'll need to flip `light` and `dark` (a little experimentation should quickly determine what looks best).

Because the navbar is styled with HTML, you'll need to `build_home_index(); init_site()` to see the effect of changing this parameter.

## Layout {#layout}

You can customise the contents of the navbar, footer, using the `navbar` and `footer` fields.
See `?build_home` for how to customise the sidebar on the homepage.
They all use a similar structure that separately defines the overall `structure` and the individual `components`.

### Navbar {#navbar-heading}

You can customise the navigation bar that appears at the top of the page with the `navbar` field.
It's made up of two pieces: `structure`, which defines the overall layout, and `components`, which defines what each piece looks like.
This organisation makes it easy to mix and match pkgdown defaults with your own customisations.

This is the default structure:

``` yaml
navbar:
  structure:
    left:  [intro, reference, articles, tutorials, news]
    right: [search, github, lightswitch]
```

It makes use of the the following built-in components:

-   `intro`: "Get Started", which links to a vignette or article with the same name as the package[^dots].
-   `reference`: if there are any `.Rd` files.
-   `articles`: if there are any vignettes or articles.
-   `tutorials`: if there any tutorials.
-   `news`: if `NEWS.md` exists.
-   `search`: the search box (see `?build_search` for more details).
-   `github`: a link to the source repository (with an icon), if it can be automatically determined from the `DESCRIPTION`.
-   `lightswitch`; a ["light switch"](#light-switch) to select light mode, dark mode, or auto mode.

Note that customising `navbar` like this comes with a downside: if pkgdown later changes the defaults, you'll have to update your `_pkgdown.yml`.

[^dots]: Note that dots (`.`) in the package name need to be replaced by hyphens (`-`) in the vignette filename to be recognized as the intro. That means for a
    package `foo.bar` the intro needs to be named `foo-bar.Rmd`.

You can use the `structure` field to reorganise the navbar without changing the default contents:

``` yaml
navbar:
  structure:
    left:  [search]
    right: [reference, articles]
```

You can use `components` to override the default content.
For example, this yaml provides a custom articles menu:

``` yaml
navbar:
 components:
   articles:
    text: Articles
    menu:
    - text: Category A
    - text: Title A1
      href: articles/a1.html
    - text: Title A2
      href: articles/a2.html
    - text: -------
    - text: "Category B"
    - text: Article B1
      href: articles/b1.html
```

Components uses the same syntax as [RMarkdown menus](https://bookdown.org/yihui/rmarkdown/rmarkdown-site.html#site-navigation).
The elements of `menu` can be:

-   Linked text (`text`, `href`, and an optional `target`). 

-   A linked icon (`icon`, `aria-label`, `href`, and an optional `target`). 
    You can find a list of available icons at [fontawesome](https://fontawesome.com/icons?d=gallery). Provide a text description of the icon in the `aria-label` field
    for screenreader users.

-   A heading (just `text`).

-   A separator (`text: ——–`).

To add a new component to the navbar, you need to modify both `structure` and `components`.
For example, the following yaml adds a new "twitter" component that appears to the left of the github icon.

``` yaml
navbar:
  structure:
    right: [search, twitter, github, lightswitch]
  components:
    twitter:
      icon: fa-twitter
      href: http://twitter.com/hadleywickham
      aria-label: Twitter
```

Finally, you can add arbitrary HTML to three locations in the navbar:

```yaml
template:
  includes:
    before_title: <!-- inserted before the package title in the header ->
    before_navbar: <!-- inserted before the navbar links -->
    after_navbar: <!-- inserted after the navbar links -->
```

These inclusions will appear on all screen sizes, and will not be collapsed into the the navbar drop down.

You can also customise the colour scheme of the navbar by using the `type` and `bg` parameters. See above for details.

### Footer

You can customise the footer with the `footer` field.
It's made up of two pieces: `structure`, which defines the overall layout, and `components`, which defines what each piece looks like.
This organisation makes it easy to mix and match the pkgdown defaults with your own customisations.

This is the default structure:

``` yaml
footer:
  structure: 
    left: developed_by
    right: built_with
```

Which uses two of the three built-in components:

-   `developed_by`: a sentence describing the main authors of the package. (See `?build_home` if you want to tweak *which* authors appear in the footer.)
-   `built_with`: a sentence advertising pkgdown.
-   `package`: the name of the package.

You can override these defaults with the `footer` field.
The example below puts the author's information on the right along with a legal disclaimer, and puts the pkgdown link on the left.

``` yaml
footer:
  structure: 
    left: pkgdown
    right: [developed_by, legal]
  components:
    legal: Provided without **any warranty**.
```

Each side is pasted into a single string (separated by `" "`) and then converted from markdown to HTML.

## Additional HTML and files

If you need to include additional HTML, you can add it in the following locations:

``` yaml
template:
  includes:
    in_header: <!-- inserted at the end of the head -->
    before_body: <!-- inserted at the beginning of the body -->
    after_body: <!-- inserted at the end of the body -->
    before_title: <!-- inserted before the package title in the header ->
    before_navbar: <!-- inserted before the navbar links -->
    after_navbar: <!-- inserted after the navbar links -->
```

You can include additional files by putting them in the right place:

-   `pkgdown/extra.css` and `pkgdown/extra.js` will be copied to the rendered site and linked from `<head>` (after the pkgdown defaults).

-   `pkgdown/extra.scss` will be added to the scss ruleset used to generate the site CSS.

-   Any files in `pkgdown/assets` will be copied to the website root directory.

-   For expert users: template files in `pkgdown/templates` will override layout templates provided by pkgdown or [template packages](#template-packages).

Use `init_site()` to update your rendered website after making changes to these files.

## Template packages {#template-packages}

To share a pkgdown style across several packages, the best workflow is to create... a package!
It can contain any of the following:

-   A configuration file in `inst/pkgdown/_pkgdown.yml`. This can be used to set (e.g.) author definitions, Bootstrap version and variables, the sidebar, footer, navbar, etc.
-   Templates in `inst/pkgdown/templates/` will override the default templates.
-   Assets in `inst/pkgdown/assets/` will be copied in to the destination directory.
    (Note these files are only copied; you'll need to reference them in your stylesheet or elsewhere in order for them to be actually used.)
-   `inst/pkgdown/extra.scss` will be added to the bslib ruleset.
    (Note that `extra.css` is not supported in templates.)

The pkgdown defaults will be overriden by these template files, which are in turn overridden by package specific settings.

Once you have created your template package `theverybest`, you need to set it as your site's theme:

``` yaml
template:
  package: theverybest
```

You then also need to make sure it's available when your site is built. Typically, you won't want to publish this package to CRAN, but you will want to publish to GitHub. Once you've done that, and assuming you're using the [usethis workflow](https://usethis.r-lib.org/reference/use_pkgdown.html), add the following line to your `DESCRIPTION`:

```yaml
Config/Needs/website: myorg/theverybest
```

This will ensure that the GitHub action will automatically install it from GitHub when building your pkgdown site.

To get some sense of how a theming package works, you can look at:

-   [tidytemplate](https://tidytemplate.tidyverse.org/) used for tidyverse and tidymodels packages;
-   [quillt](https://pkgs.rstudio.com/quillt) used for R Markdown packages;
-   [rotemplate](https://github.com/ropensci-org/rotemplate) used for rOpenSci packages.

But please note that these templates aren't suitable for use with your own package as they're all designed to give a common visual identity to a specific family of packages.

### Porting a template package

If you are updating a template package that works with pkgdown 1.0.0, create directories `inst/pkgdown/BS5/templates` and `inst/pkgdown/BS5/assets` (if you don't have any templates/assets make sure to a add dummy file to ensure that git tracks them).
The `templates` and `assets` directories directly under `inst/pkgdown` will be used by pkgdown 1.0.0 and by pkgdown 2.0.0 if `boostrap: 3`.
The directories under `inst/pkgdown/BS5/` will be used for pkgdown 2.0.0 with `boostrap: 5`.
This lets your package support both versions of Bootstrap and pkgdown.

## PR previews

Lastly, it might be useful for you to get a preview of the website in internal pull requests.
For that, you could use Netlify and GitHub Actions (or apply a similar logic to your toolset):

-   Create a new Netlify website (either from scratch by dragging and dropping a simple index.html, or by creating a site from a GitHub repository and then unlinking that repository); from the site settings get its ID to be saved as `NETLIFY_SITE_ID` in your repo secrets; from your account developer settings get a token to be saved as `NETLIFY_AUTH_TOKEN` in your repo secrets.
-   Starting from the standard pkgdown workflow `usethis::use_github_action("pkgdown")`, add some logic to build the site and deploy it to Netlify for pull requests from inside the repository, not pull requests from forks. [Example workflow](https://github.com/r-lib/pkgdown/blob/master/.github/workflows/pkgdown.yaml).

## Conclusion

In this vignette we explained how to change the theming and layout of pkgdown websites.
Further work to improve user experience will involve:

-   Working on the article (`?build_articles`) and reference indexes (`?build_reference`).
-   Writing a compelling README that explains why your package is so cool/useful/fun.
-   Improving the contents of the individual articles and reference topics 😉.
