---
title: "Introduction to pkgdown"
description: >
  Learn how to get started with the basics of pkgdown.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of pkgdown is to make it easy to make an elegant and useful package website with a minimum of work.
You can get a basic website up and running in just a couple of minutes. If you're using GitHub, we recommend setting up pkgdown and GitHub actions to automatically build and publish your site:

```{r, eval = FALSE}
# Run this once to publish your site regularly
usethis::use_pkgdown_github_pages()
```

If you are not using GitHub, you will have to run `pkgdown::build_site()` manually everytime you want to update the site.

```{r, eval = FALSE}
# Run once
# Remove docs/ from gitignore to ensure it is checked into git.
usethis::use_pkgdown()
# Run everytime you want to update your site
pkgdown::build_site()
```


While you'll get a decent website without any additional work, if you want a website that really pops, you'll need to read the rest of this vignette.
It starts by showing you how to configure pkgdown with a `_pkgdown.yml`.
You'll learn about the main components of the site (the home page, reference, articles, and news), and then how to publish and promote your site.

## Metadata

You can override pkgdown's defaults with a YAML file called `_pkgdown.yml`[^1].
The most important field is `url`, which gives the final location of the site:

[^1]: You can also put it in `pkgdown/_pkgdown.yml` if you want to keep the package root clutter-free, or in `inst/_pkgdown.yml` if you want to make it available when your package is installed. You can also use `.yaml` as the extension if desired.

``` yaml
url: https://pkgdown.r-lib.org
```

`url` is used throughout the site to generate absolute urls where they are needed.
`url` is also part of what enables auto-links to your help topics or vignettes from sites external to your package, such as from other pkgdown sites or from Quarto websites.
See `vignette("linking")` for more.

Another important option is `template`, which allows you to control the overall appearance of your site:

``` yaml
template:
  bootstrap: 5
  bootswatch: cerulean
```

You can learn more about controlling the appearance of your site in `vignette("customise")`.

## Accessibility

pkgdown's defaults work to ensure that your site is accessible to as many people as possible. But there are some accessibilty issues that only a human can solve, so make sure to also read `vignette("accessibility")` to learn about them.

## Home page

The contents of the home page are automatically generated from `index.md` or `README.md`.
pkgdown tries to put them in order, so it's possible to have a different display on GitHub and pkgdown by providing both files.
The homepage also includes a sidebar full of useful links (see `?build_home` for how these are generated and how you can customise them).

## Reference

pkgdown creates a function reference in `reference/` that includes one page for each `.Rd` help topic in `man/`.
The translation of individual help topics from Rd to HTML is generally straightforward, but there are a couple of things you should bear in mind:

-   pkgdown does its best to autolink all references to help topics and articles described in `vignette("linking")`.

-   pkgdown executes all examples, inserting the rendered results in the generated HTML files.

By default, pkgdown generates a reference index that is just an alphabetically-ordered list of functions.
The index is much more useful with human curation because functions can be grouped and described in categories.
To override the default, provide a `reference` field in `_pkgdown.yml`.

Each entry in `reference` can take one of three forms:

-   A title, defined by `title` and optional `desc` (description) fields.
-   A subtitle, defined by `subtitle` and optional `desc` (description) fields.
-   A list of topics defined by a `contents` field.

``` yaml
reference:
- title: "Connecting to Spark"
  desc: >
    Functions for installing Spark components and managing
    connections to Spark
  contents: 
  - spark_config
  - spark_connect
  - spark_disconnect
  - spark_install
  - spark_log
- title: "Reading and Writing Data"
  desc: "Functions for reading and writing Spark DataFrames."
  contents:
  - starts_with("spark_read")
  - starts_with("spark_write")
  - matches("saveload")
```

Note the use of `starts_with()` to select all functions with a common prefix.
You can also use `ends_with()` and `matches()`.
See complete details in `?build_reference`, including other topic matching helper functions.

While iterating on the reference index you might want to run `pkgdown::build_reference_index()`.
It just re-builds the index page, making it faster to quickly change `_pkgdown.yml` and see how it affects your site.

## Articles

pkgdown will automatically build all vignettes found in `vignettes/`, translating them to HTML files in `articles/`.
It is recommended to name your intro article with your package name to generate a "Get Started" page automatically.

Due to the way that pkgdown has to integrate R Markdown generated HTML with its own HTML, relatively little control is available over the output format.
You can see the details in `?build_articles`.

If you want to include an [article](https://r-pkgs.org/website.html#non-vignette-articles) on the website but not in the package (e.g., because it's large), you can use `usethis::use_article()` to set it up.

## News

If `NEWS.md` is present, it will be rendered into a single-page changelog based on markdown level headings.
pkgdown assumes your `NEWS.md` is formatted using level one headings (`#`) to specify package name and version number, and level two headings (`##`) to provide topical organization for each release.

``` markdown
# pkgdown 1.1.0

## Bug Fixes

* Lots of them

# pkgdown 1.0.0

* This is the first release of pkgdown.
```

See more suggestions for writing news bullets in the [tidyverse style guide](https://style.tidyverse.org/news.html).

See `?build_news` for more customisation options including how to:

-   Create one page for each major version and related minor versions.
-   Add release announcements to the news navbar drop-down.

## Publishing

If you use GitHub, the easiest way to build and publish your site is via GitHub actions.
Using GitHub actions automatically builds and publishes the site every time you make a change.
The easiest way to set this up is to run `usethis::use_pkgdown_github_pages()`, and if you need to customize the action, see [README.md r-lib/actions](https://github.com/r-lib/actions/tree/v2-branch/examples#build-pkgdown-site).

## Promoting

Once your finalized site is built and published on the web, you should publicize its URL in a few places:

1.  The `URL` field of your package `DESCRIPTION`, alongside a link to its source:

        URL: https://pkgdown.r-lib.org, https://github.com/r-lib/pkgdown

    (`usethis::use_pkgdown_github_pages()` does this for you.)

2.  Your repository description on GitHub.

    (`usethis::use_pkgdown_github_pages()` does this for you.)

3.  On social media (make sure to include `#rstats`).
