<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Staticdocs}
-->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  error = FALSE,
  tidy = FALSE
)
library(staticdocs)
```

# Staticdocs

## Generating a custom index

By default, staticdocs lists all functions in your package in a single list. This is not very useful for large packages, so you can use a custom `index.r` file to specify more information.

`index.r` should live in `inst/staticdocs` and should consist of calls to `sd_section()`. For example, the `index.r` for staticdocs itself looks like this:

```{r}
sd_section(
  "Settings functions",
  "These functions are used in `index.r` to configure various
    settings that staticdocs uses when rendering a package.  This is
    particularly useful for generating an index page that groups functions
    into useful categories",
  c("sd_icon", "sd_item","sd_section")
)
```

The first argument to `sd_section()` gives the name of the section. The second gives a paragraph of markdown text that will appear before the function listing, and the third argument is either a character vector of function names, or a list of `sd_teim()` objects.
