---
title: "table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a a 

```{r setup}
library(kableExtra)
library(magrittr)
```

```{r table1}
df <- dt <- mtcars[1:5, 1:6]
df %>% 
  kbl(caption = "A table ...", align = "lcccccc") %>%
  kable_styling(full_width = F,
                font_size = 13)
```


```{r table2}
df <- dt <- mtcars[1:5, 1:6]
df %>% 
  kbl(caption = "A table ...", align = "lcccccc") %>%
  kable_styling(full_width = F,
                position = "right",
                font_size = 13)
```

```{r table3}
df <- dt <- mtcars[1:5, 1:6]
df %>% 
  kbl(caption = "A table ...", align = "lcccccc") %>%
  kable_styling(full_width = F,
                position = "left",
                font_size = 13)
```
